### 2018-07-09 article/Referee responses 2.txt
minor

Referee responses 2

Comments to the Author
I thank the authors to improve the manuscript that now it is 
ready to be recommended for the publication. 
I repeat my previous comment, the lack of a (small) sample of stars 
observed with LAMOST and with normal [K/Fe] weakens the result, 
because the suspect of possible systematics remains and the 
comparison with APOGEE data is not homogeneous. 
My suggestion would be to include a sample of about 50/100 LAMOST targets 
with metallicity similar to that of the K-rich stars but 
with not enhanced K abundances, in order to quantitatively estimate 
the amount of K enrichment of these stars. 
This is my own opinion to enforce the result but I understand that this 
task is an additional and probably time-consuming effort.
The authors are free to follow or not this suggestion.

- Regrettably, undertaking the abundance estimation for LAMOST targets
- is a non-trivial endeavour, and while taking a smaller sample makes
- the problem more tractable, the reduced sample size diminishes the
- significance of the result. For this reason, we have opted to not
- include an additional sample of LAMOST stars, and have instead started
- a high-resolution spectroscopic campaign to measure detailed chemical
- abundances for a large number of normal and Mg-K stars. The detailed
- chemical abundances to be measured in that program will undoubtedly
- shed light on the origin of the Mg-K anomaly.

Additionally, I have some minor comments/suggestions.

1) Please add explicitly [Fe/H] in Table 1.

- This information is now included in Table 1, along with [K/H] and [Mg/H].

2) The link between red circles and red hexagons is not easy 
to understand at a first glance. Please better explain it 
in the caption and in Section 2.2.

- The relationship between the red circles and diamonds is now explicitly dealt with in the caption,
- as well as in the figure explanation in Section 2.2.

3) Please provide the coordinates of the targets in Table A2 
with a more significant number of digits.

- Table A2 Right Ascension and Declination coordinates are now provided to 4 decimal points (units: degrees) in Table A2.

4) Please provide the statistical significance of the 
[K/Fe]-[Fe/H] anticorrelation.

- The linear regression using the LAMOST abundances is now included in Figure 7,
- and the equation and its statistical significance quantified by r^2 is given in the caption.
- The statistical significance is fairly low, though given the large uncertainties in the abundance estimations, this is not surprising.
- The equation and r^2 value is: [K/Fe]= −0.21×[Fe/H]+0.97, with r^2 =0.37.